---
title: "Mat 125 Data Analysis"
author: "Matthew Flanders"
date: "3/30/2022"
output: pdf_document
---

```{r, echo=F, warning=F, include=F}
library(tidyverse)
library( emmeans)
library(ggridges)
load("~/Sta486Finalmtf83/data/mat125data.rda")
```


# Introduction

For the Math 125 data we were looking to answer 3 main questions:

* Did Covid-19 have a significant effect on student test scores
* Did offering a second attempt on module tests have a significant improvement in student grades that would make offering a second attempt worth the work
* Was there a significant change in student scores during covid that would indicate students cheating


# Covid-19 Effects on Grades


```{r, echo=F, warning=F}

# data becomes more normal when we drop test scores that are zero
# also working with actual tests not pre, practice, honors code or learning aids 
mat125data <- filter( mat125data, score != 0 &
                                  practice_test == 0 & 
                                  learning_aid == 0 & 
                                  honors_code == 0 & 
                                  pre_test == 0 &
                                  module_final != "F1")

# create precovid data frame, fall of 2019
f19 <- mat125data %>% filter( year == 19 & season == "fall")
f19$covid <- "fall 2019, before covid"

# create post  covid data frame, fall of 2021
f21 <- mat125data %>% filter( year == 21 & season == "fall")
f21$covid <- "fall 2021, after covid"

# bind the two data frames together for visualization
pre_post <- rbind( f19, f21)

# order the factors
pre_post <- pre_post %>% 
  mutate( covid = fct_relevel( covid,
                               "fall 2019, before covid",
                               "fall 2021, after covid"))

# plot
ggplot( pre_post, aes( x = module_final, y = score, fill = covid)) +
  geom_boxplot( alpha = 0.8) + 
  labs( title = "Pre and Post Covid In Person Module Test Scores")+
  theme(legend.position="bottom")

 ggsave("Pre and Post Covid In Person Module Test Scores.png",dpi="retina")

```

\newpage
```{r, echo=F, warning=F}
anova( lm( score ~ module_final * covid, pre_post ))

pre_post_model <- lm( score ~ module_final * covid, pre_post )

pre_post_model %>% summary()

plot( pre_post_model, which = 1:2)

```


\newpage
```{r, echo=F, warning=F}
library(emmeans)

emmeans(pre_post_model, pairwise ~ covid )

emmeans(pre_post_model, pairwise ~ covid * module_final)

```

\newpage
# Second attempt differences


```{r, echo=F, warning=F}

# build t2_attempt from mat125data using only T2 attempts and actual tests
t2_attempts <- filter( mat125data, test_attempt == "T2")


# check for duplicates
# t2_attempts[ duplicated( t2_attempts[, 1:4]), ]

# join second attempts and large data set
# added section id to join to for students who took class twice
pairwise_comp <- inner_join( t2_attempts, mat125data, by = c("Id",
                                                             "module_final",
                                                             "section_Id"))

# remove duplicate entries for tests
# mat125data has the full data set, this resulted in the previous inner join 
# matching modules with the same attempts from t2_attempts
pairwise_comp <- filter( pairwise_comp, test_attempt.x != test_attempt.y)

# duplicate check
# pairwise_comp[ duplicated( pairwise_comp[, 1:4]), ]

# create difference column
pairwise_comp$score_diff = pairwise_comp$score.x - pairwise_comp$score.y

# plot
ggplot( pairwise_comp, aes( x = module_final, y = score_diff)) +
  geom_boxplot() +
  labs( title = "Pairwise Differences Between First and Second Attempt Test Scores")+
  ylab("Difference in Test Scores")+
  xlab("Module")

ggsave("pairwise_differences.png", dpi="retina")

```
\newpage
```{r, echo=F, warning=F}
pw_model <- lm( score_diff ~ module_final, pairwise_comp )
summary(pw_model)
anova( pw_model)
plot( pw_model, which = 1:2)
```
\newpage
```{r, echo=F, warning=F}
emmeans(pw_model, pairwise ~ module_final )
```

\newpage
# Changes in remote test scores
```{r, echo = F, warning = F}

tests <- mat125data

# pre_remote
pre <- filter( tests, year == 19 & season == "fall" )
pre$remote_time_period = "Fall 2019 ( in person before covid)"

# split semester
split <- filter( tests, year == 20 & season == "spring" )
split$remote_time_period = "Spring 2020 (split)"

# remote
f20 <- filter( tests, year == 20 & season == "fall" )
s21 <- filter( tests, year == 21 & season == "spring")
f20$remote_time_period = "Fall 2020 (nau flex)"
s21$remote_time_period = "Spring 2021 (nau flex)"

# post_remote
f21 <- filter( tests, year == 21 & season == "fall" )
f21$remote_time_period = "Fall 2021 (in person after covid)"


remote_learning <- rbind(pre,split, f20, s21, f21)

remote_learning <- remote_learning %>% 
  mutate( remote_time_period = fct_relevel( remote_time_period,
                                            "Fall 2019 ( in person before covid)",
                                            "Spring 2020 (split)",
                                            "Fall 2020 (nau flex)",
                                            "Spring 2021 (nau flex)",
                                            "Fall 2021 (in person after covid)" ))

ggplot( remote_learning, aes( x = module_final, y = score, fill = remote_time_period)) +
  geom_boxplot(alpha = 0.65)+
  labs( title = "Effects of Covid-19 and Remote Learning on Module Test Scores")+theme(legend.position= "bottom")+
  guides(fill=guide_legend(nrow=3,byrow=TRUE))
  
 ggsave("Effects of Covid-19 and Remote Learning.png",dpi="retina")



```


\newpage
```{r, echo=F, warning=F}
rl_model <- lm( score ~ module_final + remote_time_period, remote_learning)
summary(rl_model)
anova( rl_model)
plot( rl_model, which = 1:2)
```


```{r, echo=F, warning=F}
emmeans(rl_model, pairwise ~  remote_time_period )
```

```{r}


# select t2 test scores and label them
t2_attempts <- pairwise_comp %>% select( c(1, 2, 4, 5))
t2_attempts <- t2_attempts %>% rename( "attempt" = 3, "score" = 4)
t2_attempts$attempt <- "Attempt 2 scores, Fall 16 - Spring 18"


# select t1 test scores and label them
t1_attempts <- pairwise_comp %>% select( c( 1, 2, 13, 14))
t1_attempts <- t1_attempts %>% rename( "attempt" = 3, "score" = 4)
t1_attempts$attempt <- "Attempt 1 scores, Fall 16 - Spring 18"

# bind first and second attempts
attempts <- rbind( t1_attempts, t2_attempts)


# length(unique( attempts$Id))
# 2250 students took tests twice

two_attempt_timeframe <- filter( mat125data, year <= 17 | ( year == 18 & season == "spring"))


# length(unique(two_attempt_timeframe$Id))
# 2852 total students during the two attempt time period


# select students who took test once by not selecting those who took the test twice
single_attempt_tests <- subset( two_attempt_timeframe, !(Id %in%  unique( attempts$Id)))
single_attempt_tests$attempt <- "Single attempt taken, Fall 16 - Spring 18"

single_attempt_tests <- select( single_attempt_tests, c(1, 2, 5, 13))

# length(unique(single_attempt_tests$Id))

no_retake <- filter( mat125data, year == 19 | ( year == 18 & season == "fall"))

no_retake$attempt = "Single attempt allowed, Fall 18 - Fall 19"

no_retake <- select( no_retake, c(1, 2, 5, 13))

df <- rbind( attempts, no_retake, single_attempt_tests)

df<- df %>% mutate( attempt = fct_relevel( attempt,
                                           "Attempt 1 scores, Fall 16 - Spring 18", 
                                           "Attempt 2 scores, Fall 16 - Spring 18",
                                           "Single attempt taken, Fall 16 - Spring 18",
                                           "Single attempt allowed, Fall 18 - Fall 19"))

ggplot( df, aes( x = score, fill = attempt))+ 
  geom_density( alpha = .4)


# ggsave("multiple and single attempt scores fall 16 to fall 19.png",dpi="retina")
# ggplot(df, aes(x = score, y = attempt, fill = attempt)) + geom_density_ridges()
```

```{r}


df <- filter( df, attempt != "Single attempt allowed, Fall 18 - Fall 19" )

ggplot( df, aes( x = score, fill = attempt))+ 
  geom_density( alpha = .4)

ggsave("multiple and single attempt scores fall 16 to spring 18.png",dpi="retina")

```




```{r}

sanity_check <- filter( mat125data, year == 19 | ( year == 18 & season == "fall"))
sum(with(sanity_check , test_attempt == "T1"))
sum(with(sanity_check , test_attempt == "T2"))

```

```{r}

single_attempt <- filter( mat125data, year == 19 | ( year == 18 & season == "fall"))

single_attempt$attempt = "single"

single_attempt <- select( single_attempt, c(1, 2, 5, 13))

df <- rbind( attempts, single_attempt)

ggplot( df, aes( x = score, fill = attempt))+ 
  geom_density( alpha = .3)


```

```{r}



t2_attempts <- pairwise_comp %>% select( c(1, 2, 4, 5, 11, 12))
t2_attempts <- t2_attempts %>% rename( "attempt" = 3, "score" = 4,
                                       "season" = 5, "year" = 6)

t1_attempts <- pairwise_comp %>% select( c( 1, 2, 13, 14, 11, 12))
t1_attempts <- t1_attempts %>% rename( "attempt" = 3, "score" = 4,
                                       "season" = 5, "year" = 6)

attempts <- rbind( t1_attempts, t2_attempts)

attempts <- filter( attempts, year < 19)

single_attempt <- filter( mat125data, year >= 19 | ( year == 18 & season == "fall"))



single_attempt$attempt = "single"

single_attempt <- select( single_attempt, c(1, 2, 5, 13, 11, 12))

df <- rbind( attempts, single_attempt)

ggplot( df, aes( x = score, fill = attempt))+ 
  geom_density( alpha = .3)+
  facet_grid( year ~ season)

# ggsave("attempts by semester.png",dpi="retina")

# ggplot(df, aes(x = score, y = year ~ season, fill = year)) + geom_density_ridges()
```

```{r}

ggplot(df, aes(x = score, y = year, fill = season)) +
  geom_density_ridges( alpha = .5)

```

